---
title: "Extract Soil Moisture data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{extract_soil_moisture_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Load the necessary Packages

```{r setup, message=F, warning=F}
library(soilmoistr)
library(iffitoR)
library(tidyverse)
library(raster)
library(stars)
library(sf)
```

# What does the function `get_sm_data` do?

- The idea of this function is to extract the soilmoisture data from a large stack of reprojected, cropped, resampled and compressed GeoTIFFs.

- **Only Landslides with a known date are considered**

- The date of the landslide, plus some days ( as specified in `days_before_window`) and minus (`days_after_window`) are then compared to the available dates of the soilmoisture-"acquisitions"

- The spatial object for which to extract the values can either be 

  * points
  * points with a bufffer 
  * polygons 
  

## When using points and no buffer

- When we are using only points (e.g. landslide initiation points) one only needs to call for example

- The extraction here is done using the `st_extract`-method for stars objects which extracts values at point locations for stars-objects

```{r exonlypoints, eval=FALSE}
res = get_sm_data(<path_to_soilmoisture_data>,
                  landsld = <landslide-object>)
```

- This will use the default values for some of the parameters:

          +  days_before_window = 5,
          +  days_after_window = 0,
          +  point_buffer = NULL,
          +  aggre_fun = NULL
          
- This mean that we look 5 days before the landslide (or any other point like object) if we find soilmoisture information

## When using points and a buffer arround them

- When you want to use a buffer around your point you can specify this in the `point_buffer`-parameter

- You directly pass the distance in units of your landsld-object. As we are mainly working in `EPSG: 32632` this should be meters

- As we are now working with a polygon, we can pass a vector with aggregation functions to the `aggre_fun`-parameter. If nothing is passed you will get back the values for all intersection cells. Look into the documentation of  [exactextracr](https://github.com/isciences/exactextractr) to know how this works in detail.

```{r expointbuffer, eval=F}
res = get_sm_data(<path_to_soilmoisture_data>,
                  landsld = <landslide-object>,
                  point_buffer = 300,
                  aggre_fun = c("mean"))

```


# Example

- lets load the landslide data from the `iffitoR`-package and subset it to only some movements of interest

```{r example, message=F, warning=F}
data = landsld[grepl("translational|rotational", landsld$second_level), ]
```

- check its dimensions

```{r dim, warning=F, message=F}
dim(data)
```

- lets create a copy and remove the column called `date` in order to see what happens
  
```{r}
copy = data.frame(data) %>% st_as_sf()
copy[["date"]] = NULL # remove the date column
```

- Now lets call the `get_sm_data`-function on that dataframe
 
```{r nodate, message=F, warning=F, error=TRUE}
res = get_sm_data(
  landsld = copy,
  path_sm = path_sm
)
```
  
- It does not work as the dataframe really needs to have a column called *date*

- Now lets use the original data to get the soilmoisture data

```{r extract, warning=F, message=F, results=FALSE}
path_sm = "/mnt/CEPH_PROJECTS/Proslide/soilmoisture/32632/"

data = data[1:50, ]

res = (get_sm_data(
  landsld = data,
  path_sm = path_sm
)) 

```

  

